Convex Surrogates via Polyhedral Losses

### 2019-05-15 notes/bo-separated.tex

\begin{lemma} \label{lemma:U-convex-closed}
  $\mathcal{U}$ is a finite set, and each element $U$ is a closed, convex polytope.
\end{lemma}
\begin{proof}
  \bo{TODO}
\end{proof}

\begin{lemma} \label{lemma:U-surject}
  There is a surjection from $\mathcal{U}$ to $\mathcal{S}$ where $U \mapsto R$ if, for some $p$, we have $\gamma(p) = U$ and $\Gamma(p) = R$.
  We write $R_U$ for the value of this map at $U$.
\end{lemma}
\begin{proof}
%  \bo{TODO!}
%  \raf{I'm not actually sure we have a bijection.  Actually, I know we don't: take hinge loss, where $\Gamma((0.9,0.1)) = \{1\}$ but $\Gamma((1,0)) = [1,\infty)$; both have $R = \{1\}$.  What we \emph{can} say is, $\mathcal{U}$ is finite, and there is a surjective map $h:\mathcal{U}\to\mathcal{S}$ where in fact $h(U) = \{r\in\R : u_r \in U\} \neq \emptyset$.  (And $u_r \in \Gamma(p) \iff r \in \gamma(p)$ of course, from embedding.)  Hopefully this is enough...}
  \bo{fixed, thanks!}
\end{proof}

@@

\begin{lemma} \label{lemma:enclose-halfspaces}
  Let $D$ be a closed, convex polytope in $\reals^d$.
  For any norm $\|\cdot\|$ and any $\epsilon > 0$, letting $B(D,\epsilon) = \{u : \min_{x \in D} \|u-x\| \leq \epsilon\}$, \bo{or open ball} there exists a convex set $D'$, the intersection of a finite number of open halfspaces, such that
    \[ D \subsetneq D' \subsetneq B(D,\epsilon) . \]
\end{lemma}
\begin{proof}
  Let $S$ be the $\epsilon$-ball in norm $\|\cdot\|$.
  Note that $B(D,\epsilon) = D + S$ where $+$ is the Minkowski sum.
  Now let $S'$ be a ball in $\|\cdot\|_1$ of nonzero radius small enough that $S' \subsetneq S$.
  (Such a ball always exists by equivalence of norms.\bo{cite})
  Because both $D$ and $S'$ are polytopes, their Minkowski sum $D' = D + S'$ is a polytope.\bo{Cite}
  Now, if $T' \subsetneq T$, then for any set $D$, the Minkowksi sum satisfies $D + T' \subsetneq D + T$, so in particular because $\emptyset \subsetneq S' \subsetneq S$, we have
    \[ D \subsetneq D' \subsetneq B(D,\epsilon) . \]
  Now we observe that because $D'$ is a polytope, we can write it as the intersection of a finite number of halfspaces, which we can make open sets.\bo{Ought to improve/reword this.}
\end{proof}

\begin{lemma} \label{lemma:thick-nonempty}
  Let $\{U_j : j \in \mathcal{J}\}$ be a finite collection of closed, convex sets with nonempty joint intersection.
  Then for all $\delta > 0$, there exists  $\epsilon > 0$ such that $\cap_j B(U_j,\epsilon) = \subseteq B(\cap_j U_j, \delta)$.
\end{lemma}
\begin{proof}
  By induction on the size of the collection.
  If the size is $1$, set $\epsilon = \delta$.
  If the size is larger than $1$, let $U_j$ be any member of the collection, let $U' = \cap_{j'\neq j} U_{j'}$, and let $C(\epsilon) = \cap_{j' \neq j} B(U_{j'},\epsilon)$.
  Let $D = U_j \cap U'$.
  We must show that $B(U_j,\epsilon) \cap C(\epsilon) \subseteq B(D,\delta)$.
  By Lemma \ref{lemma:enclose-halfspaces}, we can enclose $D$ strictly within a polytope $D'$, the intersection of a finite number of open halfspaces, which is itself strictly enclosed in $B(D,\delta)$.
  For example, if $D$ is a point, then enclose it in a hypercube, which is enclosed in the ball $B(D,\delta)$.
%  \raf{We can use the fact that the $U_j$ are all polyhedral here, so $D$ is polyhedral as well.  Now let $B_1()$ be a ball with respect to $\|\cdot\|_1$, and observe that $B_1(D,\delta')$ is therefore also polyhedral for all $\delta'$.  Using norm bounds, we can find a small enough $\delta'$ so that $B_1(D,\delta') \subseteq B(D,\delta)$.}
  We will prove that, for small enough $\epsilon$, $B(U_j,\epsilon) \cap C(\epsilon)$ is contained in $D'$.
  This implies that it is contained in $B(D,\delta)$.

@@

\begin{figure}
\caption{Illustration of a special case of the proof of Lemma \ref{lemma:thick-nonempty} where there are two sets $U_1,U_2$ and their intersection $D$ is a point. We build the polytope $D'$ inside $B(D,\delta)$. By considering each halfspace that defines $D'$, we then show that for small enough $\epsilon$, $B(U_1,\epsilon)$ and $B(U_2,\epsilon)$ do not intersect outside $D'$. So the intersection is contained in $D'$, so it is contained in $B(D,\delta)$.}
\includegraphics[width=0.24\textwidth]{figs/separated-proof-2} \hfill
\includegraphics[width=0.24\textwidth]{figs/separated-proof-3} \hfill
\includegraphics[width=0.24\textwidth]{figs/separated-proof-4} \hfill
\includegraphics[width=0.24\textwidth]{figs/separated-proof-5}
\end{figure}

>>>>

\begin{lemma} \label{lemma:U-convex-closed}
  $\mathcal{U}$ is a finite set, and each element $U$ is a closed, convex polyhedron.
\end{lemma}
\begin{proof}
  \bo{More of a sketch right now}
  Let $p$ be in the interior of the simplex.
  For any such $p$, the polyhedral expected loss $L(\cdot;p)$ projects down to the same fixed power diagram over $\reals^d$.
  The minimizer of a polyhedral loss is always the projection of some face of the polyhedron, \bo{cite} so there are a finite number of possible minimizers and each is a closed, convex polyhedron (being a member of the power diagram or an intersection of members).
  Now for $p$ not in the interior, consider each of finitely many subsets of the outcome space and restrict $p$ to have support on this subset; repeat the above arguments.
  The total number of report sets $U$ is still finite, each a closed, convex polyhedron.
\end{proof}

\begin{lemma} \label{lemma:U-surject}
  There is a surjection from $\mathcal{U}$ to $\mathcal{S}$ where $U \mapsto R$ if, for some $p$, we have $\gamma(p) = U$ and $\Gamma(p) = R$.
  We write $R_U$ for the value of this map at $U$.
\end{lemma}
\begin{proof}
  For each report $r$, write $u_r$ for its embedding point.
  By construciton, we have $u_r \in \gamma(p) \iff r \in \Gamma(p)$.
  Therefore, for each $U \in \mathcal{U}$, let $R_U = \{r : u_r \in U\}$.
  By construction, $R_U$ is nonempty because some $r$ is always optimal for any $p$, hence some $u_r \in U$ for all $U \in \mathcal{U}$.
  Furthermore, because $R_U$ are simultaneously optimal when $U$ is, i.e. when $\gamma(p) = U$, we must have that $r \in \Gamma(p) \iff u_r \in U$.
\end{proof}

@@

\begin{lemma} \label{lemma:enclose-halfspaces}
  Let $D$ be a closed, convex polyhedron in $\reals^d$.
  For any norm $\|\cdot\|$ and any $\epsilon > 0$, letting $B(D,\epsilon) = \{u : \min_{x \in D} \|u-x\| \leq \epsilon\}$, \bo{or open ball} there exists a convex set $D'$, the intersection of a finite number of open halfspaces, such that
    \[ D \subsetneq D' \subsetneq B(D,\epsilon) . \]
\end{lemma}
\begin{proof}
  Let $S$ be the $\epsilon$-ball in norm $\|\cdot\|$.
  Note that $B(D,\epsilon) = D + S$ where $+$ is the Minkowski sum.
  Now let $S'$ be a ball in $\|\cdot\|_1$ of nonzero radius small enough that $S' \subsetneq S$.
  (Such a ball always exists by equivalence of norms.\bo{cite})
  Because both $D$ and $S'$ are polyhedra, their Minkowski sum $D' = D + S'$ is a polyhedron.\bo{Cite}
  Now, if $T' \subsetneq T$, then for any set $D$, the Minkowksi sum satisfies $D + T' \subsetneq D + T$, so in particular because $\emptyset \subsetneq S' \subsetneq S$, we have
    \[ D \subsetneq D' \subsetneq B(D,\epsilon) . \]
  Now we observe that because $D'$ is a polyhedron, we can write it as the intersection of a finite number of halfspaces, which we can make open sets.\bo{Ought to improve/reword this.}
\end{proof}

\begin{lemma} \label{lemma:thick-nonempty}
  Let $\{U_j : j \in \mathcal{J}\}$ be a finite collection of closed, convex sets with nonempty joint intersection.
  Then for all $\delta > 0$, there exists  $\epsilon > 0$ such that $\cap_j B(U_j,\epsilon) = \subseteq B(\cap_j U_j, \delta)$.
\end{lemma}
\begin{proof}
  By induction on the size of the collection.
  If the size is $1$, set $\epsilon = \delta$.
  If the size is larger than $1$, let $U_j$ be any member of the collection, let $U' = \cap_{j'\neq j} U_{j'}$, and let $C(\epsilon) = \cap_{j' \neq j} B(U_{j'},\epsilon)$.
  Let $D = U_j \cap U'$.
  We must show that $B(U_j,\epsilon) \cap C(\epsilon) \subseteq B(D,\delta)$.
  By Lemma \ref{lemma:enclose-halfspaces}, we can enclose $D$ strictly within a polyhedron $D'$, the intersection of a finite number of open halfspaces, which is itself strictly enclosed in $B(D,\delta)$.
  For example, if $D$ is a point, then enclose it in a hypercube, which is enclosed in the ball $B(D,\delta)$.
%  \raf{We can use the fact that the $U_j$ are all polyhedral here, so $D$ is polyhedral as well.  Now let $B_1()$ be a ball with respect to $\|\cdot\|_1$, and observe that $B_1(D,\delta')$ is therefore also polyhedral for all $\delta'$.  Using norm bounds, we can find a small enough $\delta'$ so that $B_1(D,\delta') \subseteq B(D,\delta)$.}
  We will prove that, for small enough $\epsilon$, $B(U_j,\epsilon) \cap C(\epsilon)$ is contained in $D'$.
  This implies that it is contained in $B(D,\delta)$.

@@

\begin{figure}
\caption{Illustration of a special case of the proof of Lemma \ref{lemma:thick-nonempty} where there are two sets $U_1,U_2$ and their intersection $D$ is a point. We build the polyhedron $D'$ inside $B(D,\delta)$. By considering each halfspace that defines $D'$, we then show that for small enough $\epsilon$, $B(U_1,\epsilon)$ and $B(U_2,\epsilon)$ do not intersect outside $D'$. So the intersection is contained in $D'$, so it is contained in $B(D,\delta)$.}
\includegraphics[width=0.24\textwidth]{figs/separated-proof-2} \hfill
\includegraphics[width=0.24\textwidth]{figs/separated-proof-3} \hfill
\includegraphics[width=0.24\textwidth]{figs/separated-proof-4} \hfill
\includegraphics[width=0.24\textwidth]{figs/separated-proof-5}
\end{figure}
